---
title: "Lake Water Quality Analysis"
author: "Kaydee Barker"
date: "2/21/2022"
output: html_document
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy package/s
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
library(lubridate) #For dealing with date and time
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
#Lagos download script
#lagosne_get(dest_folder = LAGOSNE:::lagos_path(),overwrite=T)

#Load in lagos
lagos <- lagosne_load()


#Grab the lake centroid info
lake_centers <- lagos$locus

# Make an sf object 
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326)

#Grab the water quality data
nutr <- lagos$epi_nutr

#Look at column names
#names(nutr)
```

### Subset columns nutr to only keep key info that we want

```{r}
clarity_only <- nutr %>%
  dplyr::select(lagoslakeid,sampledate,chla,doc,secchi) %>%
  mutate(sampledate = as.character(sampledate) %>% ymd(.))

```


### Keep sites with at least 200 observations 

```{r}

#Look at the number of rows of dataset
#nrow(clarity_only)

chla_secchi <- clarity_only %>%
  filter(!is.na(chla),
         !is.na(secchi))

# How many observatiosn did we lose?
# nrow(clarity_only) - nrow(chla_secchi)


# Keep only the lakes with at least 200 observations of secchi and chla
chla_secchi_200 <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  mutate(count = n()) %>%
  filter(count > 200)


```


### Join water quality data to spatial data

```{r}
spatial_200 <- inner_join(spatial_lakes,chla_secchi_200 %>%
                            distinct(lagoslakeid,.keep_all=T),
                          by='lagoslakeid')

```

### Mean Chlorophyll A map

```{r}
### Take the mean chl_a and secchi by lake

means_200 <- chla_secchi_200 %>%
  # Take summary by lake id
  group_by(lagoslakeid) %>%
  # take mean chl_a per lake id
  summarize(mean_chl = mean(chla,na.rm=T),
            mean_secchi=mean(secchi,na.rm=T)) %>%
  #Get rid of NAs
  filter(!is.na(mean_chl),
         !is.na(mean_secchi)) %>%
  # Take the log base 10 of the mean_chl
  mutate(log10_mean_chl = log10(mean_chl))

#Join datasets
mean_spatial <- inner_join(spatial_lakes,means_200,
                          by='lagoslakeid') 

#Make a map
mapview(mean_spatial, zcol='log10_mean_chl', layer.name = "Mean Chlorophyll A Content")
```

**Figure 1.** Mean chlorophyl content at sites with at least 200 observations.


# Class work

## 1) What is the correlation between Secchi Disk Depth and Chlorophyll a for sites with at least 200 observations?

```{r}
ggplot(means_200) +
  geom_point(aes(mean_secchi, mean_chl)) +
  ggthemes::theme_few() + 
  xlab("Mean Secchi Disk Depth") + ylab("Mean Chlorophyll Content")

```


**Figure 2.** Chlorophyll content has a negative correlation with Secchi disk depth at sites with at least 200 observations.


### Why might this be the case? 

Secchi disks measure water clarity; the deeper the disk, the clearer the water (1). Chlorophyll content in lakes is generally a reliable marker of algae content, so that high chlorophyll values indicate high algal biomass and corresponding low water clarity (2). Additionally, chlorophyll may be used as a proxy for water quality, since high algal biomass is associated with high nutrient pollution in the process of eutrophication (2). High pollution may further decrease water clarity, so that the relationship between chlorophyll and Secchi disk depth may be expected.

1. “The Secchi Dip-in - What Is a Secchi Disk?” North American Lake Management Society (NALMS),
https://www.nalms.org/secchidipin/monitoring-methods/the-secchi-disk/what-is-a-secchi-disk/. 

2. Filazzola, A., Mahdiyan, O., Shuvo, A. et al. A database of chlorophyll and water chemistry in freshwater lakes. Sci Data 7, 310 (2020).
https://doi-org.ezproxy2.library.colostate.edu/10.1038/s41597-020-00648-2


## 2) What states have the most data? 

### 2a) Make a lagos spatial dataset that has the total number of counts per site.

```{r}

site_counts <- chla_secchi %>%
  group_by(lagoslakeid) %>%
  mutate(count = n())

lake_counts <- inner_join(site_counts, lake_centers, by= "lagoslakeid")%>% 
  dplyr::select(lagoslakeid,nhd_long,nhd_lat, count, secchi, chla)

spatial_counts <- st_as_sf(lake_counts,coords=c("nhd_long","nhd_lat"),
                          crs=4326)

```


### 2b) Join this point dataset to the us_boundaries data. 

```{r}

states <- us_states()

states_counts <- st_join(spatial_counts, states)

```


### 2c) Group by state and sum all the observations in that state and arrange that data from most to least total observations per state. 

```{r}

sum_statecount <- states_counts %>%
  group_by(state_name) %>%
  summarize(sum = sum(count)) %>%
  arrange(desc(sum)) 

sumtable <- tibble(sum_statecount)

view(sumtable)

#ggplot(data = sumtable, aes(x=state_name, y=sum, fill=state_name)) + 
#  geom_bar(stat = "identity", width = 0.3, position = "dodge") + 
#  ggthemes::theme_few() +
#  xlab("State") + ylab(expression(paste("# of Observations"))) 


```

Minnesota has the most observations. Vermont, has the next most observations, but less than half of Minnesota's observations. South Dakota has the least number of observations in the dataset.



##3 Is there a spatial pattern in Secchi disk depth for lakes with at least 200 observations?

```{r}
mapview(mean_spatial, zcol='mean_secchi', layer.name = "Mean Secchi Disk Depth")

```

**Figure 3.** Map showing mean Secchi disk depths of sites with at least 200 observations.
